### 2024-04-17 paper.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{report}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[lmargin=30mm,rmargin=30mm]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage[noblocks]{authblk}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{, }
\renewcommand*{\Authands}{, }
\renewcommand\Affilfont{\small}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{british}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Effect of the Nominal Value of Foreign Currencies in valuing a good},
  pdfauthor={Konrad Kopp},
  pdflang={en-GB},
  pdfkeywords={Behavioural Economics, Face Value Effect, Consumer Choice
Theory},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Effect of the Nominal Value of Foreign Currencies in valuing a
good}

  \author{Konrad Kopp}
            \affil{%
                  University of Oxford
              }

\date{2024-04-16}
\begin{document}
\maketitle
\begin{abstract}
Research into the effect of using foreign currencies on consumption
behaviour suggests that consumers systematically under- or overspend
based on the nominal (face) value of the foreign currency. However,
previous research is undecided on which direction this effect has, with
the two most prominent models leading to different and mutually
exclusive outcomes. This paper considers the different models and tests
their predictive and explanatory accuracy using a study. It finds that
individuals tend to underspend when the foreign currency is less
numerous than (a fraction of) the home currency and overspend when the
opposite is the case. The paper discusses which model is most accurate
at predicting and explaining this effect and considers limitations as
well as opportunities for future research.
\end{abstract}

\renewcommand*\contentsname{Table of contents}
{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{2}
\tableofcontents
}
\chapter{Introduction}\label{introduction}

Experience suggests that we treat foreign currencies differently to our
home currency and have difficulty adjusting to them. Previous research,
such as the ``money illusion'' effect (\textbf{cite?}) suggests that
individuals are biased towards the nominal (or face value) of prices.
This bias has also been explored in the context foreign currencies, with
research finding systematic effects that arise based on the numerousity
of values in a foreign currency compared to individuals' home
currencies. However, previous research is split on how and why this
effect occurs.

Raghubir and Srivastava (Raghubir and Srivastava 2002) examined the
effect of using foreign currencies on consumption behaviour and argue
that there is a systematic difference in spending behaviour based on the
exchange rate of the foreign currency to an individuals' home currency.
Specifically, they showed that when a foreign currency is a fraction of
the home currency (e.g.~1 unit in home currency is 0.2 units in foreign
currency) consumers tend to overspend in real terms relative to their
home currency. On the other hand, when the foreign currency is a
multiple of the home currency (e.g.~1 unit in home currency is 5 units
in foreign currency) consumers tend to underspend in real
terms.\footnote{For consistency, I will use the terms fraction and
  multiple currencies throughout this paper as defined here.} Raghubir
and Srivastava argue that this effect arises because individuals are
anchored on the face value of the foreign currency which influences the
exchange rate calculations from and to their home currencies. Hence,
Raghubir and Srivastava term the result of this inadequate adjustment
the ``face value effect''.

Wertenbroch, Soman and Chattopadhyay (Wertenbroch, Soman, and
Chattopadhyay 2007) consider the findings by Raghubir and Srivastava and
reiterate their conclusion that individuals inadequately calculate the
values of goods between differnet currencies, leading to systematic
differences in spending behaviour based on the exchange rate between the
foreign and home currencies. However, Wertenbroch, Soman and
Chattopadhyay argue that the process of adjustment occurs differently,
namely that individuals calculate their valuations of goods based on
salient reference points, such as budget constraints. Contrary to
standard theory, Wertenbroch, Soman and Chattopadhyay argue that
consumers care about the difference between a valuation and their budget
rather than the ratio of the two. In the context of foreign currencies,
the nominal value of this difference will diverge, leading to systematic
differences in spending behaviour. However, the direction of this bias
is the exact opposite of the one shown by Raghubir and Srivastava.
Specifically, Wertenbroch, Soman and Chattopadhyay show that consumers
underspend when the foreign currency is a fraction of the home currency
and overspend in the opposite case.

Fang (Fang 2019) further builds on top of the research by Raghubir and
Srivastava and aims to extend it to the domain of ``virtual''
currencies, which refers to non-fiat currencies that are digital in
nature and not controlled or issued by a central bank. Examples of
virtual currencies include video game (or in-game) currencies, rewards
points, airmiles, cryptocurrencies and more to the extent that these are
at least somewhat liquid. Investigating in-game currencies specifically,
Fang shows that consumer valuations systematically deviate when using
these currencies relative to the home currency. Fang's findings
correspond with those of Raghubir and Srivastava and the face value
effect.

This paper aims to build on the aforementioned research in several
important ways. Foremost, previous research is conclusive that there is
an effect that using foreign currencies has on individuals' consumption
behaviour relative to using their home currencies, but is divided on the
direction of this effect. Hence, this paper attempts to evaluate the
different models posed and use a study to test which model most
adequately predicts the perceived behaviour. Second, previous research
has only assigned minor significance to using incentive-compatible
mechanisms to elicit true preferences of participants. These mechanisms
have been established to elicit more accurate responses from
participants that are closer to their true preferences (\textbf{cite?}).
Hence, this study will use the Becker-DeGroot-Marshak (BDM) mechanism to
elicit participants' willingness to pay (WTP).{[}\^{} The BDM mechanism
will be defined and explained in more detail in the Methodology section
below.{]} Raghubir and Srivastava (Raghubir and Srivastava 2002)
expressed the concern that ``despite attempts to make the tasks as
realistic as possible, the studies reported were laboratory experiments,
and issues of generalizability when real money is on the line do arise.
For instance, when people actually exchange foreign currency they may
feel richer or poorer depending on the exchange rate, and the
differences in perceptions of wealth may affect product valuation''.
Similarly, Fang (Fang 2019) noted that implementing the BDM mechanism
was infeasible for the study he conducted but that future research
should focus on using something like it to elicit WTP. Finally, this
paper aims to extend the previous literature on the effects of using
virtual currencies specifically, due to their increasing rise in
importance when transacting or making purchasing decisions online, and
determine whether spending behaviour differs when using them.

First, I present the theoretical frameworks that have been outlined by
previous research in a standard form that makes it easy to compare and
contrast them. These frameworks are the standard model described in
consumer choice theory, the model suggested by Raghubir and Srivastava
and the model suggested by Wertenbroch, Soman and Chattopadhyay. Then, I
consider what each of these frameworks would predict in a concrete case
and show that all three models come to different and mutually exclusive
conclusions. In order to test which of these frameworks is most accurate
in predicting observed behaviour, I conduct a study in which two groups
each formulate their WTP for a specific good in an incentive compatible
way, with the only difference between the groups being the exchange rate
of the currency used to their home currency and, as a result, the
numerousity of the values of their budget and bid. Analysing the study,
I find that consumers tend to underspend when the foreign currency is a
fraction and overspend when the foreign currency is a multiple, which is
predicted by the model proposed by Wertenbroch, Soman and Chattopadhyay
(Wertenbroch, Soman, and Chattopadhyay 2007). However, due to the
limited sample size of the study, this result is not statistically
significant and does not have significant statistical power. In the
discussion, I evaluate the implications of the findings for the
theoretical frameworks outlined, consider alternative explanations for
the perceived effect and argue that the model proposed by Wertenbroch,
Soman and Chattopadhyay is the most accurate in predicting and
describing consumer behaviour when using foreign currencies. Finally, I
will consider the limitations of this paper and opportunities for
further research.

\chapter{Theoretical Framework}\label{theoretical-framework}

\section{Standard Model}\label{standard-model}

Standard consumer choice theory suggests that individuals formulate
valuations of goods in relation to reference points, such as their
budget (\textbf{cite?}). Specifically, standard theory poses that
consumers evaluate their valuation for a good as a ratio to their budget
in order to determine what proportion of their spending power is
required to purchase the good. To formulate valuations of goods in a
different currency, consumers would first determine the value of the
good in their home currency and subsequently apply the exchange rate to
get their valuation in the foreign currency. Hence, the model follows a
two step process:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Formulate valuation of good as ratio to budget in home currency
\item
  Convert the valuation into foreign currency
\end{enumerate}

Alternatively, this can be stated using the following syntax:

\[V_n = f(\dfrac{x}{M_n})\] \[V = V_n * X = V_r\]

where V is the valuation of the good in the foreign currency, \(V_n\) is
the valuation of the good in the home currency (nominal), \(V_r\) is the
valuation of the good in the foreign currency, \(M_n\) is the budget in
the home currency, X is the exchange rate and \(f(x)\) is some function
that uses a valuation relative to the budget in order to compute
\(V_n\). Note that \(V_r\) is simply \(V_n\) multiplied by the exchange
rate, whereas V need not always be identical to \(V_r\).

One assumption that is implicit in this model is that a consumer will
always correctly adjust their valuation ratio based on the exact
exchange rate. However, this might not always be the most rational thing
to do. For example, when exchange rates fluctuate it might not be worth
the effort of a consumer to always use the most up-to-date exchange
rate, but rather to use some approximate value. Further, when an
individual needs to do these exchange rate calculations in their head,
it might not be worth the mental effort to calculate the amount in the
foreign currency perfectly, but an approximation might suffice. For
example, it would be rational to multiply an amount in home currency by
1.2 if the real exchange rate is 1.2113248 and the effort required is
significant, such as when attempting to calculate it in one's hand. By
loosening these assumptions, the standard model is able to account for
small deviations from a consumer's real valuation of a good in a foreign
currency.

\section{Face Value Effect Model}\label{face-value-effect-model}

Raghubir and Srivastava (Raghubir and Srivastava 2002) put forth a
different model for how consumers formulate their valuation of a good in
a foreign currency based on their studies conducted. Specifically, they
pose that a consumer will first formulate a valuation of a good in their
home currency, before converting that amount into the foreign currency
using the exchange rate. However, they argue that this calculation is
biased by the nominal value of the good in the home currency (in cases
where a consumer does not determine their valuation but observes a
price, this calculation would be biased by the nominal value of the good
in the foreign currency). This model may be formulated as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Formulate valuation of good in home currency
\item
  Convert the valuation into foreign currency
\item
  Adjust using the nominal value of good in home currency
\end{enumerate}

Alternatively, this can be stated using the following syntax:

\[V = α(V_n) + (1 - α)V_r\]

where α is a weighting parameter such that \(α ∈[0,1]\). Importantly,
Raghubir and Srivastava (Raghubir and Srivastava 2002) argue that
\(α > 0\) and that V is hence not equal to \(V_r\) but is biased towards
\(V_n\).

\section{Perceived Value of Money
Model}\label{perceived-value-of-money-model}

Wertenbroch, Soman and Chattopadhyay (Wertenbroch, Soman, and
Chattopadhyay 2007) build on top of Raghubir and Srivastava, but put
forth a different model that they argue is more representative of the
behaviour of actual consumers. Specifically, they agree with the
standard model that valuations are made in relation to salient reference
points, such as a consumers' budget, and argue that the model put forth
by Raghubir and Srivastava fails to account for this. However, they
disagree with the standard model that valuations use a ratio of price to
budget, but instead argue that consumers think in terms of differences.
Hence, instead of considering what fraction of their purchasing power is
required to buy a good, consumers consider how much of their budget is
left over. However, Wertenbroch, Soman and Chattopadhyay argue that
consumers inadequately adjust this difference to the exchange rate and
are thus biased when considering goods in different currencies since the
nominal differences between price and budget will differ when
considering different currencies. This model can be formulated as
follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Formulate valuation of good as difference to budget in home currency
\item
  Convert the valuation into foreign currency
\item
  Adjust using the difference to budget in the foreign currency
\end{enumerate}

Alternatively, this can be stated using the following syntax:

\[V = α(M_r - V_r) + (1 - α)V_r\]

where \(M_r\) is the budget in the foreign currency (real). Importantly,
Wertenbroch, Soman and Chattopadhyay (Wertenbroch, Soman, and
Chattopadhyay 2007) argue that \(α > 0\) and that V is hence not equal
to \(V_r\) but is biased towards \(M_r - V_r\).

\section{Predictions}\label{predictions}

The three different models laid out above would each generate a
different prediction in the following scenario, considering fraction and
multiple exchange rates (the latter in brackets): a consumer values a
good at £25 and needs to express this valuation in BlueCoin, a fictional
currency. The exchange rate of Pounds to BlueCoin is 1 BlueCoin = £5 (1
BlueCoin = £0.2) and their budget is 10 (250) BlueCoin. The models above
would make the following predictions:

\begin{itemize}
\tightlist
\item
  Standard model: \(V_r = 5 (125) BlueCoin\)
\item
  Face value effect model: \(V_r > 5 (< 125) BlueCoin\)
\item
  Perceived value of money model: \(V_r < 5 (> 125) BlueCoin\)
\end{itemize}

As a result:

\begin{itemize}
\tightlist
\item
  Standard model: \(V_rf = V_rm\)
\item
  Face value effect model: \(V_rf > V_rm\)
\item
  Perceived value of money model: \(V_rf < V_rm\)
\end{itemize}

where \(V_rf\) is the valuation of the good in the foreign currency for
the fraction group and \(V_rm\) is the valuation of the good in the
foreign currency for the multiple group.

In other words, the standard model would expect the valuations to be the
same across both exchange rates, the face value effect model would
expect the consumer to overspend when presented with the fraction
exchange rate (and vice-versa) and the perceived value of money model
would expect the consumer to underspend when presented with the fraction
exchange rate (and vice-versa).

\chapter{Method}\label{method}

To test which model is most accurate in predicting and explaining
consumption behaviour, I conducted an online study. There were 41
participants in the study, most of them current undergraduate students
at the University of Oxford, but some participants were also recent
graduates and postgraduate students. The study took place online, using
Qualtrics, and students were incentivised to participate by taking part
in a lottery, which will be explained in further detail below.

\section{Study}\label{study}

The study itself was a survey split into four parts: consent,
introduction, auction and follow-up Questions. The entire experiment
instructions can be found in the Experiment Instructions section of the
Appendix. The first section asked participants to read through and sign
the consent form in order to ensure that they were informed about the
study and what data would be collected. In the introductory section, the
BDM mechanism was explained to participants and they were asked to
complete four question that tested their understanding of the mechanism.
The aim of this section was to ensure that participants properly
understood how the mechanism worked in order to elicit their true WTP
during the main part of the study.

The auction was the main part of the study, with the aim of eliciting
the participants' WTP on a specific item and comparing these across
different exchange rates used. In this section, participants were
randomly split into two groups with almost-identical instructions. The
participants were instructed to bid on an item shown below, a tabletop
airhockey table, using the virtual currency ``BlueCoin''. The auction
was explained to follow the BDM mechanism and it was also stated that
three participants of the experiment would be randomly selected for
their auction to be simulated and to receive the outcome (either their
budget in Pounds or the item and remaining budget). The only difference
between the groups was the exchange rate between BlueCoin and the
British Pound (£) and the endowment that the participants received in
BlueCoin. The first group received an endowment of 10 BlueCoin where 1
BlueCoin equals £5 and the second group received an endowment of 250
BlueCoin where 1 BlueCoin equals £0.2. To complete the auction,
participants were asked to respond with their bid, in BlueCoin.

The final part of the experiment consisted of optional follow-up
questions that were used to determine if there are any trends in bidding
behaviour based on some characteristics of the participants. The
questions asked were about whether they had previously lived in a
different country than the UK, how often they tend to travel outside of
the UK and what their typical expenditure is in a given month. The
intention behind the first two questions is to use them as a proxy for
experience dealing with foreign currencies, where presumably people that
travel frequently or have lived in multiple countries are more
experienced with currency conversion calculations. The final question
was asked to be able to determine whether there were any systematic
differences in WTP based on a participants normal expenditure.

\section{Auction Design}\label{auction-design}

Since the aim of the study was to determine whether there were any
systematic differences in WTP between consumers when the only differing
factor is the numerousity of the currency used, I decided to not include
a control group that bid on the item using Pounds, which was the design
of some of the experiments run by Raghubir and Srivastava (Raghubir and
Srivastava 2002). This allowed for a simpler design and more power given
the same sample size. Further, unlike the studies run by Raghubir and
Srivastava (Raghubir and Srivastava 2002) and Wertenbroch, Soman and
Chattopadhyay (Wertenbroch, Soman, and Chattopadhyay 2007), I did not
use a fiat currency or mix of fiat and virtual currencies, but used only
one virtual currency for both groups. On the one hand, this
simplification was due to a focus of wether the effects described in
previous papers also held in similar ways for virtual currencies. On the
other hand, it also allowed me to preclude any biases or previous
experience that participants might have had towards or with certain
currencies. Raghubir and Srivastava (Raghubir and Srivastava 2002)
acknowledge that ``the face value effect is due to the accessibility and
perceptual salience of the face value of the foreign currency \ldots{}
{[}which{]} is likely to depend on the extent to which an individual has
the opportunity or the time available to process exchange rate
information and/or has experience in using a particular foreign
currency''. Further, Alter and Oppenheimer (Alter and Oppenheimer 2008)
show that people use ``familiarity and fluency'' when valuing goods in
different currencies. To bias participants the least, I chose the
relatively nondescript name for a fictional, virtual currency: BlueCoin.

Another important decision in the design of the experiment was which
item to select that a majority of participants would have a non-zero
WTP. This is important since if the majority or even the entirety of
participants were to bid zero on the item, then I would only learn that
this was the participants' WTP for that good, but not whether there are
any systematic differences in the WTP when using different currencies.
Hence, I selected a tabletop airhockey table, something that few enough
people have so that they would not not want another one, but enough
people want with a non-zero WTP, especially among a student population.
This turned out to be a good choice, since only six participants of the
experiment stated a WTP of 0.

In designing the main part of the experiment, I chose to conduct an
auction, specifically using the BDM mechanism, in order to have an
incentive-compatible method of eliciting the participants' WTP. The
BDM's incentive compatibility in eliciting WTP is a well-established
phenomenon that is superior to using participants' stated preferences
(\textbf{cite?}). The BDM is essentially a sealed-bid, second price
auction that is, in its' most common variant, played against a computer
drawing a bid from a known distribution. If the players' bid is greater
than or equal to the computers' bid, they win and pay the computers'
bid, otherwise they lose. As mentioned above, the studies conducted by
Raghubir and Srivastava (Raghubir and Srivastava 2002), Wertenbroch,
Soman and Chattopadhyay (Wertenbroch, Soman, and Chattopadhyay 2007)
(with the exception of one permutation of one of the studies) and Fang
(Fang 2019) did not implement incentive-compatible methods of eliciting
the participants' preferences, making this experiment an important
addition to these previous studies in examining whether the effects
described still hold when a more truthful way of eliciting WTP is used.
In order to incentivise the participants to participate, I ran a lottery
that paid out three randomly-chosen participants based on their actual
choices made in the auction and by simulating the BDM mechanism using a
random number generator. Due to budget constraints, I was only able to
play this lottery for three people, but given that the sample size of
the experiment was relatively small, the potential to receive one's
outcome is likely to have helped in eliciting the participants' true
WTP.

Finally, given the budget, I was able to give participants an endowment
of £50 each, which was around double of what the item costs to buy.
Assuming that the item is priced roughly around the average WTP (which
turned out to be correct), this endowment was chosen to leave enough
room for people with higher WTP to make their bid and to capture this in
the data. This is important for the goals of the experiment, since if I
had, for example, chosen an endowment of £20, the majority of
participants might have bid their maximum amount, hindering my ability
to meaningfully compare WTP across the two groups. I chose relatively
simple exchange rates of 5 and \(1/5\) so that the exchange rate
calculations would be relatively easy to make, allowing me to be more
confident that differences between groups were not due to the mental
effort required to do calculations but due to something else. Raghubir
and Srivastava (Raghubir and Srivastava 2002) acknowledge that ``the
reliance on face value may be a function of the ease with which the
foreign money can be converted. \ldots{} Future research should examine
the asymmetric and nonlinear nature of this effect''. Since the goal of
this study is to investigate that an effect exists that is not solely
based on a high difficulty or large mental effort required when doing
specific, difficult calculations, I aimed to make the calculations as
easy as possible in order to isolate the effect of the exchange rate and
nominal value of the foreign currency. Further, since the study was
conducted online, participants were easily able to and allowed to use
calculators for any part.

\section{Limitations}\label{limitations}

In setting up the experiment, I incorrectly used the BDM mechanism.
Instead of explaining it as a sealed-bid second-price auction, the
experiment instructions explained it as a sealed-bid first-price
auction. Unlike in a second-price auction, the participants of a first
price auction do not have a (weakly) dominant strategy to bid their true
valuations, in fact there generally is no dominant strategy in such an
auction (\textbf{cite?}). This is true even in this auction in which
participants knew that they were bidding against a computer agent who
would randomly draw a bid from a uniform distribution from 0-50, with an
expected bid of 25. Indeed, the optimal strategy by participants was to
choose their bid b by maximising their payoff p:

\[\frac{b}{50} * u(x = 1, 50 - b) + (1 - \frac{b}{50}) * u(x = 0, 50)\]

where u(x, m) is their utility function, x is a binary variable where x
= 1 signifies owning the good (and vice-versa) and m = 50 is their
budget (in Pounds). As is obvious, the value of b depends on the
participants' utility function. Hence, while players in this auction
need not necessarily have the optimal strategy of telling their true
valuation, sufficient randomisation should ensure that the utility
functions of agents in each group should be evenly distributed. This
would entail that we can nonetheless compare bids across both groups.

Another limitation of the study is the low sample size of 41
participants, which entails lower statistical power and thus a lower
likelihood of detecting a true positive when an effect actually exists.
I will examine this in further detail in the next section, conducting a
power analysis and estimating how big the sample size should have been
to achieve an adequate level of power.

\chapter{Results}\label{results}

There were 41 participants, 21 in the first group and 20 in the second.
To conduct the data analysis, I have normalised the bids across both
groups by converting them into pounds and computed the following summary
statistics:

\begin{verbatim}
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
   0.00   10.00   20.00   19.74   32.00   50.00      20 
\end{verbatim}

\begin{verbatim}
[1] 15.49001
\end{verbatim}

\begin{verbatim}
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
   0.00   16.50   27.50   25.57   35.25   50.00      21 
\end{verbatim}

\begin{verbatim}
[1] 14.8083
\end{verbatim}

Further, I plotted the following boxplot, in order to visually see the
difference between the groups:

\includegraphics{paper_files/figure-pdf/unnamed-chunk-2-1.pdf}

As is obvious from the summary statistics and the boxplot, there is a
difference between both groups, namely that the bids in the multiple
group seem to be higher on average than those in the fraction group. To
investigate this hypothesis, I conducted a t-test of a difference in
means between the groups, specifically a Welch Two Sample t-test. The
null hypothesis is that the difference in means is equal to 0 (or,
alternatively, that the means are the same) and the alternative
hypothesis is that this difference is not equal to 0 (or, alternatively,
that the means are not the same):

\[h_0: µ_1 = µ_2\] \[h_1: µ_1 ≠ µ_2\]

Hence, I conducted the t-test:

\begin{verbatim}

    Welch Two Sample t-test

data:  experiment_data$group_fraction_bid_in_pounds and experiment_data$group_multiple_bid_in_pounds
t = -1.2323, df = 38.999, p-value = 0.2252
alternative hypothesis: true difference in means is not equal to 0
95 percent confidence interval:
 -15.401901   3.740092
sample estimates:
mean of x mean of y 
  19.7381   25.5690 
\end{verbatim}

The t-value of -1.2323 suggests that there is a difference in the means
between the groups and, since it is negative, that the mean of the
fraction group is lower. However, the p-value is 0.2252, which is quite
high. If we take the typical significance level of \(α = 0.05\) we find
that the p-value is higher than this value, and thus fail to reject the
null hypothesis at the 5\% significance level. Further, the 95\%
confidence interval is \([-15.401901, 3.740092]\) which includes 0,
providing further evidence that the difference in means between the
groups is not statistically significant. Thus, we must conclude that
while there seems to be a difference in means based on the sample data,
this difference is not statistically significant for this sample size.

To investigate this further, I have conducted a power analysis in order
to determine the power of the t-test, which refers to the likelihood of
the test detecting a true positive when an effect actually exists
(\textbf{cite?}). High power would indicate that this likelihood is high
and vice versa. To conduct the analysis, we first need to find the
effect size, or Cohen's D, and then implement the power analysis.
Further, I will also conduct an analysis that investigates how big the
sample size should have been to get a power of 80\%, which is a
generally accepted threshold for significant power.

\begin{verbatim}

     t test power calculation

             n1 = 21
             n2 = 20
              d = 0.3845806
      sig.level = 0.05
          power = 0.2246363
    alternative = two.sided
\end{verbatim}

\begin{verbatim}

     Two-sample t test power calculation

              n = 107.1047
              d = 0.3845806
      sig.level = 0.05
          power = 0.8
    alternative = two.sided

NOTE: n is number in *each* group
\end{verbatim}

Based on the first test with the same significance level of 0.05, we can
see that the power is 0.2246363 or 22\%, which means that there is only
a 22\% likelihood of detecting a true positive if an effect actually
exists. In other words, the likelihood of detecting a true difference in
means between the two groups, when this difference exists, is only 22\%,
which is far below the usual value of 80\%. The second test takes in
this power of 80\% and calculates the required sample size at the
observed effect size. The result is \(n = 107\) where the sample size is
\(2n\) or 214. Hence, using a sample size of 214 and observing the same
effect, this t-test would yield a power of 80\% at the 5\% significance
level.

Finally, I conducted a regression of the demographic values collected in
the fourth part of the experiment on the bids of each group. This is
done in order to determine whether there is any correlation between the
demographic variables and the bid amount, which would suggest that
randomisation did not occur correctly, which could have been exacerbated
by the small sample size. The regression yields the following outcomes:

\begin{verbatim}

Call:
lm(formula = group_fraction_bid_in_pounds ~ lived_abroad + travel_frequency + 
    expenditure, data = experiment_data)

Residuals:
    Min      1Q  Median      3Q     Max 
-20.744 -10.085   0.000   7.481  33.333

Coefficients: (2 not defined because of singularities)
                                           Estimate Std. Error t value Pr(>|t|)
(Intercept)                                  16.667     11.144   1.496    0.169
lived_abroadNo                               16.883     34.290   0.492    0.634
lived_abroadYes                               4.651     28.453   0.163    0.874
travel_frequencyLess than once per 5 years   12.768     31.087   0.411    0.691
travel_frequencyOnce per 1-2 years           -4.952     27.679  -0.179    0.862
travel_frequencyOnce per 2-5 months           5.085     23.987   0.212    0.837
travel_frequencyOnce per 3-5 years          -18.005     42.079  -0.428    0.679
travel_frequencyOnce per 6-11 months         -6.488     24.395  -0.266    0.796
travel_frequencyOnce per month or more           NA         NA      NA       NA
expenditure£251-£400                         -6.317     17.686  -0.357    0.729
expenditure£401-£550                         -5.544     19.548  -0.284    0.783
expenditure£551-£700                         -8.884     20.685  -0.429    0.678
expenditureAbove £1000                      -13.549     37.738  -0.359    0.728
expenditureBelow £250                            NA         NA      NA       NA

Residual standard error: 19.3 on 9 degrees of freedom
  (20 observations deleted due to missingness)
Multiple R-squared:  0.3013,    Adjusted R-squared:  -0.5528 
F-statistic: 0.3527 on 11 and 9 DF,  p-value: 0.9466
\end{verbatim}

\begin{verbatim}

Call:
lm(formula = group_multiple_bid_in_pounds ~ lived_abroad + travel_frequency + 
    expenditure, data = experiment_data)

Residuals:
    Min      1Q  Median      3Q     Max 
-15.667  -4.550   0.000   5.383  17.200

Coefficients: (2 not defined because of singularities)
                                       Estimate Std. Error t value Pr(>|t|)  
(Intercept)                              20.000     13.715   1.458   0.1788  
lived_abroadNo                           -0.600     17.250  -0.035   0.9730  
lived_abroadYes                          17.600     17.250   1.020   0.3342  
travel_frequencyOnce per 1-2 years        4.600     21.843   0.211   0.8379  
travel_frequencyOnce per 2-5 months      12.400     17.250   0.719   0.4905  
travel_frequencyOnce per 6-11 months      7.267     19.717   0.369   0.7210  
travel_frequencyOnce per month or more       NA         NA      NA       NA  
expenditure£251-£400                     -5.000     18.025  -0.277   0.7877  
expenditure£401-£550                    -17.200     15.801  -1.089   0.3046  
expenditure£551-£700                    -43.487     21.461  -2.026   0.0734 .
expenditure£701-£850                     -8.400     17.250  -0.487   0.6379  
expenditureAbove £1000                  -25.000     19.396  -1.289   0.2296  
expenditureBelow £250                        NA         NA      NA       NA  
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

Residual standard error: 13.72 on 9 degrees of freedom
  (21 observations deleted due to missingness)
Multiple R-squared:  0.5937,    Adjusted R-squared:  0.1422 
F-statistic: 1.315 on 10 and 9 DF,  p-value: 0.3456
\end{verbatim}

Based on this regression, there are no statistically significant
correlations at the 5\% significance level or lower. However, there is
one correlation that is significant at the 10\% significance level,
namely of \texttt{expenditure£551-£700} on the multiple group. The
coefficient of -43.487 suggests that people whose usual expenditure is
between £551 and £700 per month bid lower than the group overall.
However, given that this is only significant at the 10\% level and not
at any lower ones, it is likely that this correlation is due to the low
sample size rather than it representing a true correlation between these
variables.

\chapter{Discussion}\label{discussion}

The study conducted found that the fraction group bid less on average
than the multiple group. The difference in the means across the two
groups is around £6, or over 10\% of the budget, signifying that this
difference occurs due to a systematic difference in how each group
calculated their value in the foreign currency BlueCoin rather than
random chance. The only difference between the groups was the exchange
rate of BlueCoin to the Pound and, as a result, the numerousity of the
nominal values of the budget and the bids. This would suggest that this
difference in bids occurs based on whether the foreign currency is a
fraction or multiple of the home currency.

Above, I outlined the three models for how consumers value goods in a
foreign currency that have been suggested by previous research. The
standard model would have predicted that both groups would on average
bid the same, the model proposed by Raghubir and Srivastava (Raghubir
and Srivastava 2002) would have predicted that the fraction group would
on average bid more than the multiple group and the model proposed by
Wertenbroch, Soman and Chattopadhyay (Wertenbroch, Soman, and
Chattopadhyay 2007) would have predicted the opposite, namely that the
fraction group would on average bid less than the multiple group. Hence,
the third model would have made the most accurate prediction for the
study conducted.

Wertenbroch, Soman and Chattopadhyay (Wertenbroch, Soman, and
Chattopadhyay 2007) argue that this systematic difference in valuations
and spending behaviour occurs because consumers are biased by the
nominal difference between their budget and the product valuation in the
foreign currency and thus anchor on this value when converting amounts
between their home currency and foreign currency. In the context of the
study, this would entail that the fraction group calculated the
difference between their budget and their bid in BlueCoin and adjusted
their bid downwards based on it, since this difference will be small
(\textless{} 10) in nominal terms. The multiple group, on the other
hand, would have calculated this difference and received a larger amount
of surplus for the same product valuation in real terms (25 times as
large as the fraction group). Hence, the multiple group would have
adjusted their bid upwards, leading to a higher average bid in real (and
nominal) terms.

However, there might be alternative explanations that would be
consistent with one or more of the other models outlined above. One
explanation would be that consumers are biased by only the nominal value
of their budget in the foreign currency and not by the difference
between their budget and their valuation of a product in the foreign
currency. While this alternative hypothesis would be hard to prove, it
would not change the prediction for how the difference in product
valuations across groups that use either a fraction or multiple
currency. Indeed, it would predict the exact same result and is thus
consistent with the model proposed by Wertenbroch, Soman and
Chattopadhyay, even though it differs in explaining the internal work
that individuals would do to come to that conclusion.

Another explanation might be that people are simply worse at dividing
than multiplying (or the reverse), which could explain the perceived
differences between the groups. On this explanation, even when consumers
have the same valuation in real terms, one group will systematically
fail to accurately convert this value into the foreign currency.
However, this explanation is unlikely for several reasons. First, it
assumes that a significant amount of people are significantly worse at
performing one of the operations across all numbers. From experience, it
seems that when considering most numbers, one of these operations is
more difficult to do but it is not always the same one. Secondly, this
assumes that people are unaware enough of their difficulties (even when
they are very significant) that they do not use a calculator even when
one is easily accessible, such as when completing the survey for this
study online. For these reasons, this explanation seems unlikely to be
accurate.

A third contending explanation could be that individuals simply prefer
to use whole numbers over fractions or decimals, leading consumers to
make less granular valuation choices when considering a currency with a
low face value. Because of this, there can be differences in groups of
individuals if, for example, many participants in the same group choose
to round down. For example, it could have been the case that the average
valuation of the good in the fraction group was £24 but if a significant
amount of participants rounded this down to the nearest whole number in
BlueCoin (4), then this average valuation would now be 20. This
explanation can be supported empirically, given that only 3 bids out of
41 were decimal numbers. One might also claim that this behaviour of
preferring whole numbers is rational, given that it requires less mental
effort, especially when multiplying and dividing these numbers. Further
research is required to determine whether this hypothesis could explain
the observed behaviour, either entirely or in conjunction with one of
the models proposed above.

The former two of these alternative explanations seem unlikely, but the
latter seems like it could be accurate at least to some extent. Without
further research, however, it seems like the model proposed by
Wertenbroch, Soman and Chattopadhyay is the most accurate in predicting
and explaining the perceived behaviour when formulating valuations of
goods in different foreign currencies.

\section{Limitations, Theoretical Implications and Future
Research}\label{limitations-theoretical-implications-and-future-research}

The main limitation of the study conducted is that due to small sample
size, the results are not statistically significant at the 5\% level and
the power of the t-test for a difference in means is only 22\%. Hence,
even though the observations line up with the model proposed by
Wertenbroch, Soman and Chattopadhyay, it could be the case that this
result is due to random chance. In repeating this study, it would be
useful to get a sample size of at least 200 participants to roughly
achieve a power of 80\% and likely also statistical significance in
conducting the t-test. The second major limitation is that the BDM
mechanism was implemented incorrectly, entailing that there was no
dominant strategy for participants to reveal their true valuations on
the good. In a redo, this mistake could be fixed relatively easily by
adjusting the experiment instructions.

Apart from the limitations, the major theoretical implication of this
paper is the finding that the model proposed by Wertenbroch, Soman and
Chattopadhyay (Wertenbroch, Soman, and Chattopadhyay 2007) seems to be
most accurate in predicting and explaining consumer behaviour when
eliciting a WTP on a good in a foreign currncy. In line with these
authors, this conclusion contradicts the findings of the ``face value
effect'' by Raghubir and Srivastava (Raghubir and Srivastava 2002) and
subsequently by Fang (Fang 2019). Further, this conclusion also implies
that, contrary to traditional consumption theory, individuals formulate
their WTP in relation to the difference between their budget and the
valuation of a good rather than its' ratio.

Future research could build on this study and its' findings in several
important ways: first, it would make sense to conduct a similar or
identical study with the BDM mechanism implemented correctly and a
greater sample size. This would allow for greater certainty that the
effect observed is statistically significant and reproducible. Further,
it would be interesting to investigate the third alternative hypothesis
formulated above in order to gain a better understanding into the exact
process that leads individuals to systematically under- or overspend
when using foreign currencies. It might, however, be difficult to
conduct an adequate study to accurately test that claim.

\chapter{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-Alter2008}
Alter, Adam L., and Daniel M. Oppenheimer. 2008. {`Easy on the Mind,
Easy on the Wallet: The Roles of Familiarity and Processing Fluency in
Valuation Judgments'}. \emph{Psychonomic Bulletin {\&} Review} 15 (5):
985--90. \url{https://doi.org/10.3758/PBR.15.5.985}.

\bibitem[\citeproctext]{ref-Fang2019}
Fang, Justin. 2019. {`In-Game Currency Design and Consumer Spending
Behavior'}. Bachelor\textquotesingle s Thesis, Ross School of Business,
University of Michigan.
\url{https://deepblue.lib.umich.edu/bitstream/handle/2027.42/155343/Justin\%20Fang_BA\%20480\%20Written\%20Report.pdf}.

\bibitem[\citeproctext]{ref-Raghubir2002}
Raghubir, Priya, and Joydeep Srivastava. 2002. {`{Effect of Face Value
on Product Valuation in Foreign Currencies}'}. \emph{Journal of Consumer
Research} 29 (3): 335--47. \url{https://doi.org/10.1086/344430}.

\bibitem[\citeproctext]{ref-Wertenbroch2007}
Wertenbroch, Klaus, Dilip Soman, and Amitava Chattopadhyay. 2007. {`{On
the Perceived Value of Money: The Reference Dependence of Currency
Numerosity Effects}'}. \emph{Journal of Consumer Research} 34 (1):
1--10. \url{https://doi.org/10.1086/513041}.

\end{CSLReferences}

\chapter{Appendix}\label{appendix}

\section{Experiment Instructions}\label{experiment-instructions}

\begin{figure}[H]

{\centering \includegraphics{experiment_instructions/consent.png}

}

\caption{Consent}

\end{figure}%%
\begin{figure}[H]

{\centering \includegraphics{experiment_instructions/instructions.png}

}

\caption{Instructions}

\end{figure}%%
\begin{figure}[H]

{\centering \includegraphics{experiment_instructions/auction_fraction.png}

}

\caption{Auction: Fraction Group}

\end{figure}%%
\begin{figure}[H]

{\centering \includegraphics{experiment_instructions/auction_multiple.png}

}

\caption{Auction: Multiple Group}

\end{figure}%%
\begin{figure}[H]

{\centering \includegraphics{experiment_instructions/follow_up.png}

}

\caption{Follow up}

\end{figure}%

\section{Source Code}\label{source-code}

The full source code of data analysis is follows:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Load the required libraries}
\FunctionTok{library}\NormalTok{(pwr)}
\FunctionTok{library}\NormalTok{(lsr)}

\CommentTok{\# Load the experiment data}
\NormalTok{experiment\_data }\OtherTok{\textless{}{-}}
  \FunctionTok{read.csv}\NormalTok{(}\StringTok{"virtual{-}currencies{-}paper/data/experiment\_data\_formatted.csv"}\NormalTok{)}

\CommentTok{\# Summary statistics for the bid columns}
\FunctionTok{summary}\NormalTok{(experiment\_data}\SpecialCharTok{$}\NormalTok{group\_fraction\_bid\_in\_pounds)}
\FunctionTok{sd}\NormalTok{(experiment\_data}\SpecialCharTok{$}\NormalTok{group\_fraction\_bid\_in\_pounds, }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{)}

\FunctionTok{summary}\NormalTok{(experiment\_data}\SpecialCharTok{$}\NormalTok{group\_multiple\_bid\_in\_pounds)}
\FunctionTok{sd}\NormalTok{(experiment\_data}\SpecialCharTok{$}\NormalTok{group\_multiple\_bid\_in\_pounds, }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{)}

\CommentTok{\# Plot a boxplot for the bid columns}
\FunctionTok{boxplot}\NormalTok{(}
\NormalTok{  experiment\_data}\SpecialCharTok{$}\NormalTok{group\_fraction\_bid\_in\_pounds,}
\NormalTok{  experiment\_data}\SpecialCharTok{$}\NormalTok{group\_multiple\_bid\_in\_pounds,}
  \AttributeTok{main =} \StringTok{"Boxplot of fraction and multiple group bids (in Pounds)"}\NormalTok{,}
  \AttributeTok{names =} \FunctionTok{c}\NormalTok{(}\StringTok{"Group Fraction"}\NormalTok{, }\StringTok{"Group Multiple"}\NormalTok{),}
  \AttributeTok{xlab =} \StringTok{"Variables"}\NormalTok{,}
  \AttributeTok{ylab =} \StringTok{"Values"}
\NormalTok{)}

\CommentTok{\# Perform a t{-}test to compare the means of the two groups}
\FunctionTok{t.test}\NormalTok{(}
\NormalTok{  experiment\_data}\SpecialCharTok{$}\NormalTok{group\_fraction\_bid\_in\_pounds,}
\NormalTok{  experiment\_data}\SpecialCharTok{$}\NormalTok{group\_multiple\_bid\_in\_pounds}
\NormalTok{)}

\CommentTok{\# Conduct a regression analysis to examine the relationship between}
\CommentTok{\# the bid amounts and the demographic variables}
\NormalTok{model\_group\_fraction }\OtherTok{\textless{}{-}}
  \FunctionTok{lm}\NormalTok{(}
\NormalTok{    group\_fraction\_bid\_in\_pounds }\SpecialCharTok{\textasciitilde{}}\NormalTok{ lived\_abroad }\SpecialCharTok{+}
\NormalTok{      travel\_frequency }\SpecialCharTok{+}\NormalTok{ expenditure,}
    \AttributeTok{data =}\NormalTok{ experiment\_data}
\NormalTok{  )}
\FunctionTok{summary}\NormalTok{(model\_group\_fraction)}

\NormalTok{model\_group\_multiple }\OtherTok{\textless{}{-}}
  \FunctionTok{lm}\NormalTok{(}
\NormalTok{    group\_multiple\_bid\_in\_pounds }\SpecialCharTok{\textasciitilde{}}\NormalTok{ lived\_abroad }\SpecialCharTok{+}
\NormalTok{      travel\_frequency }\SpecialCharTok{+}\NormalTok{ expenditure,}
    \AttributeTok{data =}\NormalTok{ experiment\_data}
\NormalTok{  )}
\FunctionTok{summary}\NormalTok{(model\_group\_multiple)}

\CommentTok{\# Calculate the effect size (Cohens D)}
\NormalTok{effect\_size }\OtherTok{\textless{}{-}} \FunctionTok{cohensD}\NormalTok{(}
\NormalTok{  experiment\_data}\SpecialCharTok{$}\NormalTok{group\_fraction\_bid\_in\_pounds,}
\NormalTok{  experiment\_data}\SpecialCharTok{$}\NormalTok{group\_multiple\_bid\_in\_pounds}
\NormalTok{)}

\CommentTok{\# Find the power of the t test conducted above}
\FunctionTok{pwr.t2n.test}\NormalTok{(}\AttributeTok{n1 =} \DecValTok{21}\NormalTok{, }\AttributeTok{n2 =} \DecValTok{20}\NormalTok{, }\AttributeTok{d =}\NormalTok{ effect\_size, }\AttributeTok{sig.level =} \FloatTok{0.05}\NormalTok{)}

\CommentTok{\# Find the sample size required to achieve a power of 0.8 (80\%)}
\FunctionTok{pwr.t.test}\NormalTok{(}\AttributeTok{power =} \FloatTok{0.8}\NormalTok{, }\AttributeTok{d =}\NormalTok{ effect\_size, }\AttributeTok{sig.level =} \FloatTok{0.05}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\end{document}